---
title: "SISBID Day 2, Afternoon"
author: "Kim"
date: "July 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
```


Subsetting with NAs
```{r, chunk2}
x = c(1,4,5,6,NA,7)
x > 5 # This preserves NA

y <- x %in% c(5,7)
y
```

Tables:

Check missingness
```{r, chunk3}
table(x, useNA = "ifany")
#This will add another column with the number of NAs
unique(x)
table(x)
table(x, useNA = "always")
#This will always include a column for NA even if there aren't any

```



Tables with factors
```{r, chunk4}
fac = factor(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             levels = 1:4) #0 gets dropped because you did not specify 0 as a level
tab = table(fac)
tab

fac1 = factor(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             levels = 0:4) #0 gets dropped because you did not specify 0 as a level
tab1 = table(fac1)
tab1
```

Create a two way table
```{r, chunk5}
tab2 <- table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
             c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3), 
              useNA = "always")
tab2

tab3 <- table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), 
             c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3), 
              useNA = "always", dnn = c("First", "Second"))
tab3 #DNN will add labels to the tables
```

Recoding to and from missing: is.na
```{r, chunk6}
ages = c(23,21,44,32,57,65,-999,54)
range(ages) # Range will show you the range!

ages[ages == -999] = NA #Recode -999 to missing
range(ages)
range(ages,na.rm=TRUE) # Take a look at non-missing

#To missing
is.na(ages)
ages[is.na(ages)] = -999
ages
```

Download data and take a look: 
```{r, chunk7}
Sal = read.csv("../data/Baltimore_City_Employee_Salaries_FY2014.csv",
               as.is = TRUE)
str(Sal)
```

Recode, etc. - see slides and lab?

Paste
```{r, chunk8}
paste("Visit", 1:5, sep = "_")
paste("Visit", 1:5, sep = "_", collapse = " ")
 #Collapse puts it all into one vector
paste("To", "is going be the ", "we go to the store!", sep = "day ")
  #This can be useful for putting a bunch of vars together for regression
```

Using stringr and regular expressions:
  RE are ways to search for specific strings
  http://www.regular-expressions.info/reference.html 
```{r, chunk9}
x <- c("I really", "like writing", "R code programs")
y <- strsplit(x, split = " ") # returns a list
y
```

```{r, chunk10}
mycities <- c("Seattle", "SF","LA", "Chicago")
y <- strsplit(mycities, split = " ") # returns a list
y


```

sapply in the slides -> applying the function to each part of the list? I still don't understand this.

grep
```{r,chunk11}
grep("Rawlings", Sal$Name) #Will find names with Rawling in them
Sal[grep("Rawlings", Sal$Name),] #return values in base R

Sal %>% filter(str_detect(Name, "Rawlings")) #dplyr
```

Sub and gsub
```{r, chunk12}

Sal$AnnualSalary <- as.numeric(gsub(pattern = "$", replacement="", 
                              Sal$AnnualSalary, fixed=TRUE))
#gsub takes $, replaces it with nothing in annual salary and need fixed 
#because $ has a value in regular expressions
Sal <- Sal[order(Sal$AnnualSalary, decreasing=TRUE), ] 
Sal[1:5, c("Name", "AnnualSalary", "JobTitle")]
```

With dplyr
```{r, chunk13}
dplyr_sal = Sal
dplyr_sal = dplyr_sal %>% mutate(   #mutate creates a new column but giving it the same name
  AnnualSalary = AnnualSalary %>%
    str_replace(
      fixed("$"), 
      "") %>% #Piping
    as.numeric) %>%
  arrange(desc(AnnualSalary))
check_Sal = Sal
rownames(check_Sal) = NULL
all.equal(check_Sal, dplyr_sal)
```